Study of Intra- and Inter-Job Interference with Different Job allocations

### 2015-10-15 submission.tex
list key finds in introduction

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Indication of multiple jobs running concurrently with different allocations. Each job is represented by a specific color. a) shows the effect of partition-based allocation, which reduce the inter-job interference. b) shows non-contiguous allocation, which introduces both intra and inter-job interferences. }
   \label{fig: overview}
\end{figure}

One of the most prominent  problems is network contention among concurrently running jobs. The network sharing among concurrently running jobs causes communication variability that results in jobs running slower \cite{abhinav-sc13}. The delay of running jobs increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}.

This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource. The batch scheduler is responsible for assigning job the amount of nodes it requires. On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used. In the partition based allocation, used in Blue Gene series systems, he scheduler assigns each user job a compact and contiguous set of computing nodes. This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention caused by concurrently running jobs. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible. The other strategy is non-contiguous allocation, used by Cray XT/XE series, in which free nodes are assigned to user jobs whether they are contiguous or not. Non-contiguous allocation eliminates internal and external fragmentation as seen in partition-based systems, thereby leading to high system utilization. However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.

%\textcolor{red}{
%Partition-based allocation achieves good job performance by sacrificing system utilization, whereas non-contiguous allocation can result in better system utilization but could severely degrade performance of user jobs. What is needed is a more flexible strategy that provides allocations that minimize intra/inter-job contention but allow greater system utilization. In this work, we propose a guideline for future batch scheduling systems, a smart way to allocate resources with the awareness of jobs communication patterns.
%
%Most existing batch schedulers take so few initiative because it been hold under the obscurantism policy. Whether it use partition-based or non-contiguous policy, means it chooses the benefit of either system or job over the other. Some propose compromised policies like buddy system\cite{buddy system paper} still make the allocation for job blindly without knowing job's communication information. 
%
%Although HPC applications complexity continue to grow, people could get more knowledge about the applications communication information with the help from various parallel application profiling tools, such as TAU\cite{tau}, MPIP\cite{mpip} and DUMPI\cite{sst}. Job's communication trace and performance analysis can be get by using such tool at runtime. 
%
%The off-line study of job's communication traces could be helpful to develop better allocation policies for batch scheduler to make on-line scheduling/allocating decisions. This is because the communication pattern of parallel applications on HPC system can be summarized into a couple of categories\cite{roth}. The deep understanding of one application can be helpful to all other applications that conforms to the same communication pattern. Job's communication traces study would be extreme helpful to leadership computing facility too. The set of applications running on those leadership computing facilities is usually stable, and the workloads from those HPC systems is usually has great repetitiveness. We analyzed the workload on Mira, an IBM Blue/Gene Q system at Argonne Leadership Computing Facility, found that most jobs require 2k nodes and about 14 particular jobs consist of the majority its workload\ref{fig: repetitiveness of Mira}. In other word, Mira accommodates mostly some particular 14 jobs and most jobs have the same size.
%}

We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both partition-based allocation and non-contiguous allocation. Such flexible mechanism would make allocation for jobs based on their communication patterns. With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job. Rather than assign each job a compact partition-based allocation or non-contiguous allocation, the ``neighborhood communication" consists of such tightly coupled processes should get a compact allocation. The new allocation mechanism only needs to guarantee the locality of the subset of the nodes where those ``neighborhood communication" would reside.

In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Torus networks have been extensively used in the current generation of supercomputers because of their linear scaling on per-node cost and competitive communication performance. On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the “Tofu” system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns. We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations. We analyze the intra- and inter-job interference by simulating applications running exclusively or concurrently with different allocations.

\textbf{replace this by summarizing a couple of key findings!!} According to results of our comprehensive exploration, making allocation with consideration about job's communication pattern is a very promising allocation strategy.

The rest of this paper organized as follows. Section \ref{sec:application study} gives a detailed study of the three applications from DOE Design Forward Project. Section \ref{sec:codes} talks about CODEs as research vehicle for our work. Section \ref{sec:config study} shows the performance analysis of three applications on different torus networks. Section \ref{sec:interference} provides detailed analysis about the intra/inter-job interference between three applications on torus network. Section \ref{sec:discussion} talks about the idea of smart allocation strategy with job communication pattern awareness. Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.

@@

2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus. 2D mesh is a very common allocation shape in torus network for both partition-based and non-contiguous allocation strategies. For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}. When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh. The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.

@@

%We simulate three application running concurrently with different allocation to analysis the interference between them. First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively. For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks. CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others. The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than partition-based system. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

@@

Based on our comprehensive study, we can get following observations 
\begin{enumerate}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keeps relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefers compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the neighborhood communication of the application will result in better performance. 
\end{enumerate}

@@

The advantage of making allocation with consideration about job's communication pattern is obvious. First, compared with partition-based policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application. On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern. On the contrary, the small compact node set (allocation unit) provided for the application's  ``local communities"  will preserve the communication locality in the maximum extent.

>>>>

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/goodallocation}
        \caption{Contiguous}
        \label{fig:overview_sub1}
    \end{subfigure}%
    \hspace{1em}%
    \begin{subfigure}[t]{0.2\textwidth}
        \centering
        \includegraphics[height=1.2in]{figs/badallocation}
        \caption{Non-contiguous}
        \label{fig:overview_sub2}
    \end{subfigure}%
   \caption{Multiple jobs running concurrently with different allocations. Each job is represented by a specific color. a) shows the effect of contiguous allocation, which reduce the inter-job interference. b) shows non-contiguous allocation, which introduces both intra and inter-job interferences. }
   \label{fig:overview}
\end{figure}

Users make job submissions to the HPC system with specification about job's required number of nodes and expected runtime. Batch scheduler will pick several jobs and dispatch them to the system for running. These jobs run concurrently and utilize the system in a space sharing way. One of the most prominent problems is network contention among concurrently running jobs. The network sharing among concurrently running jobs causes communication variability that results in jobs running slower \cite{abhinav-sc13}. The delay of running jobs also increases the queueing time of the following submitted jobs, thus leading to low system throughput and utilization \cite{jose-ipdps15}. This adverse effect of network sharing can be mitigated by providing jobs with isolated allocations and exclusive network resource.

On the widely used torus-connected HPC systems \cite{bgq}\cite{tofu}\cite{titan}, two allocation strategies are commonly used. In the contiguous allocation, the scheduler assigns each user job a compact and contiguous set of computing nodes, as shown in Figure \ref{fig:overview_sub1}. The partition-based allocation used in Blue Gene series systems is one of this kind \cite{zhou-ipdps}. This strategy favors application performance by providing application with isolated partitions and exclusive network connections, thus reducing network contention from concurrently running jobs. However, this strategy can cause internal fragmentation (when more nodes are allocated to a job than it requests) and external fragmentation (when sufficient nodes are available for a request, but they can not be allocated contiguously), therefore leading to lower system utilization that is otherwise possible. The other strategy is non-contiguous allocation, used by Cray XT/XE series \cite{carl-cug}, in which free nodes are assigned to jobs whether they are contiguous or not. Figure \ref{fig:overview_sub2} shows the effect of non-contiguous allocation. Non-contiguous allocation eliminates internal and external fragmentation as seen in contiguous allocation systems, thereby leading to high system utilization. However, it introduces other problems, for example scattering application processes all over the system, causing both inter-job and intra-job contention. The non-contiguous node allocation can significantly reduce job performance, especially for communication-intensive ones.

We envision that future HPC will adopt a flexible job allocation mechanism which combines the best of both contiguous and non-contiguous allocation. Such flexible mechanism would make allocation for jobs based on their communication patterns. With detail analysis of application's communication patterns, it can be identified which processes are tightly connected and conduct intensive communication within each job. Rather than assign each job a whole chunk of compact contiguous allocation, only the ``neighborhood communication" consists of such tightly coupled processes need to get a compact allocation. The new allocation mechanism only needs to guarantee the compactness of the subset of the nodes where those ``neighborhood communication" would reside.

In this work, we focus on an in-depth analysis of intra- and inter-job communication interferences with  different job allocations on torus-connected HPC systems. Torus networks have been extensively used in the current generation of supercomputers because of their good scalability and competitive communication performance. On the recent Top500 list, 6 of the top 10 supercomputers use a high-radix torus-interconnected network \cite{top500}. The current generation of IBM Blue Gene/Q (BG/Q) supercomputer, such as Mira at Argonne Nation Laboratory and Sequoia at  Lawrence Livermore National Laboratory, has its nodes connected in a 5D torus network \cite{bgq}. The K computer from Japan uses the ``Tofu” system, which is a 6D mesh/torus topology \cite{tofu}. Titan, a Cray XK7 supercomputer located at the Oak Ridge Leadership Computing Facility (OLCF), has nodes connected in a 3D torus within the compute partition \cite{titan}. Although the analysis are based on torus network, the idea conveyed from this work is applicable to networks with different topologies.

We selected three signature applications from DOE Design Forward Project\cite{designforwardwebpage} as examples to conduct detailed study about their communication patterns. We use a sophisticated simulation toolkit named CODES, Co-Design of Multi-layer Exascale Storage Architecture\cite{Jason-2011}, from Argonne National Laboratory as a research vehicle to evaluate the performance of these applications with various allocations. We analyze the intra- and inter-job interference by simulating these applications running exclusively and concurrently with different allocations. The major finds and insights of this work are the following:

\begin{itemize}
    \item We study the communication behaviors and identify the communication patterns of three representative HPC applications.
    \item We conduct detailed analysis about the intra- and inter-job interference by running three applications with different allocations. And we find the relation between application's communication pattern and the cause of interference.
    \item We show a novel and flexible way to make allocation with awareness of job's communication pattern, which  is every effective to alleviate the interference between concurrently running jobs. 
\end{itemize}

We believe the finds and insights presented in this work would be very useful for the design of future HPC batch job scheduler and resource management module.

The rest of this paper is organized as follows. Section \ref{sec:application study} gives a detailed study of the three representative applications from DOE Design Forward Project. Section \ref{sec:codes} talks about CODEs as research vehicle for our work. Section \ref{sec:config study} shows the performance analysis of three applications on different torus networks. Section \ref{sec:interference} provides detailed analysis about the intra- and inter-job interference between three applications on torus network with different allocations. Section \ref{sec:discussion} talks about the idea of a novel and flexible allocation strategy with job communication pattern awareness. Section \ref{sec:related_work} talks about the related work in the area and conclusion will be presented in Section \ref{sec:conclusion}.

@@

2D mesh, shown as blue in Figure \ref{fig:cont_sub1}, can be cut out from a single layer of the 3D torus. 2D mesh is a very common allocation shape in torus network for both contiguous and non-contiguous allocation strategies. For example, Cray Application Level Placement Scheduler indexes the nodes in torus network into a list and make allocation by simply works off the this list \cite{carl-cug}. When the list is obtained by sorting the nodes based on their spacial coordinates in the torus, allocation made off from this list will be 2D mesh. The IBM Blue/Gene Q supercomputer Mira at Argonne Leadership Computing Facility also allow its allocation partition configured into mesh \cite{zhou-ipdps}.

@@

%We simulate three application running concurrently with different allocation to analysis the interference between them. First, we assign each application with contiguous allocation side by side, and compare their performance with each application running exclusively. For AMG, sharing the network with other application won't affect its performance, since all the data transfer are between neighboring ranks. CrystalRouter and MultiGrid will suffer prolonged communication time when sharing network with others. The results are shown in Figure \ref{fig: concurrent-cont-time-box}.
Inter-job interference has been identified as one of the major culprits for application's performance variability \cite{abhinav-sc13}\cite{skinner}\cite{rosenthal}. Inter-job interference is a more prominent issue for system adopted non-contiguous allocation policy than system with contiguous allocation. Application communication times have been demonstrated to vary from 36\% faster to 69\% slower due to job interference when running on non-contiguous allocation \cite{abhinav-sc13}.

@@

Based on our comprehensive study, we can get following observations:
\begin{itemize}
    \item The compact allocation may not guarantee the best performance for every application. 
    \item The performance of application conforms to Nearest Neighbor communication pattern keeps relatively stable with different shape allocations, while application conforms to ``Many-to-Many" communication pattern prefers compact allocation. 
    \item The allocation unit size should be decided according to application's communication pattern. Unit size that big enough to preserve the neighborhood communication of the application will result in better performance. 
\end{itemize}

@@

The advantage of making allocation with consideration about job's communication pattern is obvious. First, compared with contiguous allocation policy, the new allocation strategy is more flexible, it doesn't require the system to provide a big contiguous partition that can accommodate the whole application, just small set of compact nodes enough for all the "local community" in the application. On the other hand, the new allocation strategy won't destroy the locality of application's communication pattern. On the contrary, the small compact node set (allocation unit) provided for the application's  ``local communities"  will preserve the communication locality in the maximum extent.
